### 2020-12-10 doc/manuscript/Dow_manuscript_Nature.Rmd
minor
include DOYtiming figure instead of climwin
#41

put DOY timing figure back in main manuscript

removed climwin altogether, as its repetitive with existing SI figures

Co-Authored-By: Cameron Dow <55721987+camerondow35@users.noreply.github.com>

#Figure numbers
schematic_summary=1 #illustrates parameters
climwin = 2 # probably stays?
pheno_Tsensitivity = 3 
tree_rings = 4  #stays, hopefully adding a lot more sites

@@

climwin_caption <- "Figure 3. Landscapes of relationships between the day of year on which 25% of annual growth is achieved ($DOY_{25}$) and temperature in prior weeks for ring- and diffuse-porous trees at SCBI and Harvard Forest. Shown are matrices of linear coefficients of first-order linear regressions between temperature and $DOY_{25}$, where Window Open and Window Close indicate number of weeks prior to $DOY_{25}$ (ring-porous: May 5 at SCBI, May 13 at HF; diffuse-porous: June 4 at SCBI, June 14 at HF). Black circles indicate the critical $T_{max}$ window (ring-porous:  March 22- April 9 at SCBI, April 2 - May 07 at HF; diffuse-porous: Feb. 19- May 21 at SCBI, March 19 - May 07 at HF)."

@@

@@

#SI figure numbers
growth_curves_all = "S1"  
all_climwin_SI_figs = " S2-S5"
  climwin_rp_DOY25_SCBI = "S2"
  climwin_dp_DOY25_SCBI = "S3"
  climwin_rp_DOY25_HARV = "S4"
  climwin_dp_DOY25_HARV = "S5"
DOYtiming = "S6"
  DOYtiming_SCBI = "S6a"
  DOYtiming_Harvard = "S6b"
tree_rings_TMIN = "S7"

#SI figure captions
growth_curves_all_caption  <- "Growth trajectories for ring- and diffuse-porous trees, as both relative and cumulative fractions of total annual growth. Each line represents one year's growth for a given tree, fit with McMahon model. "

climwin_rp_DOY25_SCBI_short_caption <- "Full climwin output for $DOY_{25}$ for ring porous species at SCBI."
climwin_dp_DOY25_SCBI_short_caption <- "Full climwin output for $DOY_{25}$ for diffuse porous species at SCBI."
climwin_rp_DOY25_HARV_short_caption <- "Full climwin output for $DOY_{25}$ for ring porous species at Harvard Forest."
climwin_dp_DOY25_HARV_short_caption <- "Full climwin output for $DOY_{25}$ for diffuse porous species at Harvard Forest."
DOYtiming_caption <- "Day of year where growth milestones were achieved at (a) the Smithonian Conservation Biology Institute and (b) Harvard Forest. Each line represents a single year of ring-porous (dotted line) or diffuse-porous (solid line) trees at each site. Mean temperature was calculated for each wood-type/site combination over the respective critical $T_{max}$ window, then turned into a ratio and assigned a color on a gradient where the coldest year in the sample is blue and the warmest is red."
tree_rings_TMIN_short_caption <- "Climate sensitivity of annual growth, as derived from tree-rings, for dominant ring- and diffuse- porous species at the Smithsonian Conservation Biology Institute and Harvard Forest. Shown are Pearson correlations between ring-width index and and monthly Tmin. Species codes are as given in Table 2.(THIS FIGURE WILL PROBABLY BE PROBABLY MODIFIED SIGNFICANTLY TO INCLUDE LOTS MORE SITES.)"

@@

For the majority of weekly time windows from Jan. 1 to $DOY_{25}$, there was a negative or neutral relationship between $T_{max}$ and $DOY_{25}$, meaning that warmer temperatures shifted growth earlier in the season (Figs. `r climwin`, `r all_climwin_SI_figs`). 
This held true for both ring- and diffuse-porous trees at both sites, *albiet* with less consistent patterns at Harvard Forest, which was sampled for fewer years. 
The strongest relationships, including but not limited to the optimal time windows (Table `r mean_growth_parameters_table`), occurred for time windows closing $\le$ 5 weeks prior to $DOY_{25}$.

```{r climwin, out.width="80%", echo=FALSE, fig.align="center", fig.cap= climwin_caption}
# Created in RScripts/additional_figures.R
knitr::include_graphics("tables_figures/climwin_figure.png")
```

The correlation of $DOY_{25}$ to $T_{min}$ was less consistent than that of $T_{max}$ (Figs. `r climwin`, `r all_climwin_SI_figs`). 
At SCBI, patterns were quite similar to those described above for $T_{max}$, with similar timing and strengths of correlations across the two temperature variables (Figs. `r climwin_rp_DOY25_SCBI` - `r climwin_dp_DOY25_SCBI`).
In contrast, at Harvard Forest, where we had only six years of data, *climwin* selected windows, both > 5 weeks prior to $DOY_{25}$, had positive correlations between $DOY_{25}$ and $T_{min}$.
However, for both ring- and diffuse- porous species, average $T_{min}$ of the selected period was below freezing (<0° C) in all years, suggesting that these are sporadic correlations that are unlikely to be biologically meaningful.

@@

For both ring- and diffuse- porous species at both sites, warmer spring $T_{max}$ resulted in earlier achievement of growth milestones (Fig. `r schematic_summary`, Figs. `r pheno_Tsensitivity`).
Specifically, $DOY_{25}$, $DOY_{50}$, and $DOY_{g.max}$ at SCBI and $DOY_{25}$, $DOY_{50}$, $DOY_{75}$, and $DOY_{g.max}$ at Harvard Forest all decreased significantly with increasing $T_{max}$ during the critical temperature window (Figs. `r schematic_summary`, `r pheno_Tsensitivity`).

@@

@@

Although the number of years analyzed here is limited (10 years at SCBI, 6 years at HF), the signal is clear and consistent: for both species groups at both sites, warmer temperatures during virtually any part of the spring led to significantly earlier achievement of growth milestones (Figs. `r schematic_summary`-`r pheno_Tsensitivity`).
This indicates that the effect of $T_{max}$ on the timing of growth ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$,and $DOY_{g_{max}}$) is not subtle; 
indeed, growth milestones advanced by 1-8 days per degree Celsius during the critical temperature window.
This is consistent with previous studies showing that ... *(find/cite some statistics on how rapidly leaf phenology or start of C sequestration advance with T)*

@@

Our findings contribute to a broader picture of the influence of spring temperatures on forest ecosystem function. 
Warm spring temperatures result in earlier leaf-out (fig., REFS), an earlier start to net C sequestration (REFS), and--as shown here--earlier initiation of stem growth (Figs. `r climwin`- `r pheno_Tsensitivity`).
Yet, an earlier start to the growing season does not imply a longer growing season (Figs. `r schematic_summary`, `r pheno_Tsensitivity`) [@zani_increased_2020].
Rather, growth finishes sooner-- presumably because of *C sink limitation* [@zani_increased_2020].

@@

new Science paper: @zani_increased_2020

>>>>

#Figure numbers
schematic_summary=1 #illustrates parameters
pheno_Tsensitivity = 2 
DOYtiming = 3
  DOYtiming_SCBI = "3a"
  DOYtiming_Harvard = "3b"
tree_rings = 4  #stays, hopefully adding a lot more sites

@@

@@

DOYtiming_caption <- "Day of year where growth milestones were achieved at (a) the Smithonian Conservation Biology Institute and (b) Harvard Forest. Each line represents a single year of ring-porous (dotted line) or diffuse-porous (solid line) trees at each site. Mean temperature was calculated for each wood-type/site combination over the respective critical $T_{max}$ window, then turned into a ratio and assigned a color on a gradient where the coldest year in the sample is blue and the warmest is red. (WE WILL ADD LEAF PHENOLOGY HERE)"

@@

#SI figure numbers
growth_curves_all = "S1"  
all_climwin_SI_figs = " S2-S5"
  climwin_rp_DOY25_SCBI = "S2"
  climwin_dp_DOY25_SCBI = "S3"
  climwin_rp_DOY25_HARV = "S4"
  climwin_dp_DOY25_HARV = "S5"

tree_rings_TMIN = "S7"

#SI figure captions
growth_curves_all_caption  <- "Growth trajectories for ring- and diffuse-porous trees, as both relative and cumulative fractions of total annual growth. Each line represents one year's growth for a given tree, fit with McMahon model. "
climwin_rp_DOY25_SCBI_short_caption <- "Full climwin output for $DOY_{25}$ for ring porous species at SCBI."
climwin_dp_DOY25_SCBI_short_caption <- "Full climwin output for $DOY_{25}$ for diffuse porous species at SCBI."
climwin_rp_DOY25_HARV_short_caption <- "Full climwin output for $DOY_{25}$ for ring porous species at Harvard Forest."
climwin_dp_DOY25_HARV_short_caption <- "Full climwin output for $DOY_{25}$ for diffuse porous species at Harvard Forest."

tree_rings_TMIN_short_caption <- "Climate sensitivity of annual growth, as derived from tree-rings, for dominant ring- and diffuse- porous species at the Smithsonian Conservation Biology Institute and Harvard Forest. Shown are Pearson correlations between ring-width index and and monthly Tmin. Species codes are as given in Table 2.(THIS FIGURE WILL PROBABLY BE PROBABLY MODIFIED SIGNFICANTLY TO INCLUDE LOTS MORE SITES.)"

@@

For the majority of weekly time windows from Jan. 1 to $DOY_{25}$, there was a negative or neutral relationship between $T_{max}$ and $DOY_{25}$, meaning that warmer temperatures shifted growth earlier in the season (Figs. `r all_climwin_SI_figs`). 
This held true for both ring- and diffuse-porous trees at both sites, *albiet* with less consistent patterns at Harvard Forest, which was sampled for fewer years. 
The strongest relationships, including but not limited to the optimal time windows (Table `r mean_growth_parameters_table`), occurred for time windows closing $\le$ 5 weeks prior to $DOY_{25}$.

The correlation of $DOY_{25}$ to $T_{min}$ was less consistent than that of $T_{max}$ (Figs. `r all_climwin_SI_figs`). 
At SCBI, patterns were quite similar to those described above for $T_{max}$, with similar timing and strengths of correlations across the two temperature variables (Figs. `r climwin_rp_DOY25_SCBI` - `r climwin_dp_DOY25_SCBI`).
In contrast, at Harvard Forest, where we had only six years of data, *climwin* selected windows, both > 5 weeks prior to $DOY_{25}$, had positive correlations between $DOY_{25}$ and $T_{min}$.
However, for both ring- and diffuse- porous species, average $T_{min}$ of the selected period was below freezing (<0° C) in all years, suggesting that these are sporadic correlations that are unlikely to be biologically meaningful.

@@

For both ring- and diffuse- porous species at both sites, warmer spring $T_{max}$ resulted in earlier achievement of growth milestones (Fig. `r schematic_summary`, Figs. `r pheno_Tsensitivity`).
Specifically, $DOY_{25}$, $DOY_{50}$, and $DOY_{g.max}$ at SCBI and $DOY_{25}$, $DOY_{50}$, $DOY_{75}$, and $DOY_{g.max}$ at Harvard Forest all decreased significantly with increasing $T_{max}$ during the critical temperature window (Figs. `r schematic_summary`, `r pheno_Tsensitivity`, `r DOYtiming`).

@@

```{r DOYtiming, out.width="100%", echo=FALSE, fig.align="center", fig.cap= DOYtiming_caption}
# Created in RScripts/additional_figures.R
knitr::include_graphics("tables_figures/DOYtiming_allyears.png")
```

@@

Although the number of years analyzed here is limited (10 years at SCBI, 6 years at HF), the signal is clear and consistent: for both species groups at both sites, warmer temperatures during virtually any part of the spring led to significantly earlier achievement of growth milestones (Figs. `r schematic_summary`-`r `r DOYtiming``).
This indicates that the effect of $T_{max}$ on the timing of growth ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$,and $DOY_{g_{max}}$) is not subtle; 
indeed, growth milestones advanced by 1-8 days per degree Celsius during the critical temperature window.
This is consistent with previous studies showing that ... *(find/cite some statistics on how rapidly leaf phenology or start of C sequestration advance with T)*

@@

Our findings contribute to a broader picture of the influence of spring temperatures on forest ecosystem function. 
Warm spring temperatures result in earlier leaf-out (fig., REFS), an earlier start to net C sequestration (REFS), and--as shown here--earlier initiation of stem growth (Figs. `r schematic_summary`- `r DOYtiming`).
Yet, an earlier start to the growing season does not imply a longer growing season (Figs. `r schematic_summary`, `r pheno_Tsensitivity`) [@zani_increased_2020].
Rather, growth finishes sooner-- presumably because of *C sink limitation* [@zani_increased_2020].

@@
